---
title: "R Assignment Template"
author: "Amelia"
date: '2022-05-10'
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Introduction 

Here you will answer some fundamental questions like, but are not limited to: 
1. What is the purpose of this assignment?
2. What data will you be using? Where is it from, how was it collected, when etc. 
3. Are there any importatntn details about the data?
4. Why is this important?

```{r}
# code may include summaries, head/tail, na, imputation, etc. 
```


#### Methods 

1. What type of statistical test are you using?
2. Why does it make sense for this data?
3. What do you hope to gain out from this methodology?

```{r}
# run tests 
```


#### Results 

1. What are the numeric results of the test?
2. Did everything go how you expected?
3. What do the results mean?
4. Is anything significant?
5. This is where you would but graphs

```{r}
# report results 
```

#### Conclusion 

1. What general conclusions can you make from the results?
2. What does this mean?
3. How is this significant?
4. How did the testing go?
5. What shoud be done different/going forward?