---
title: "MSDS 660 Week 3 Homework Assignment"
author: "Jeremy Beard"
date: '2022-05-21'
output:
  pdf_document: default
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Introduction 

In this week's assignment, we will be exploring a dataset found somewhere on the internet. We will be loading and cleaning data, performing regression analysis, looking at Variance Inflation Factor (VIF) values, performing stepAIC analysis, creating boxplots and histograms, and more. This assignment will be a great introduction to Multiple Linear Regression, not just single variable linear regression. 

The dataset I chose to explore was analyzing air quality data, specifically the concentration of CO2 in the atmosphere in one area of Italy. There are some null values in the data so some data cleaning was implemented.

This assignment is important because it teaches the fundamentals of Multiple Linear Regression analysis in R. It builds upon skills already built up previously and deals heavily in model comparison and comprehending model performance parameters.

```{r}
# Let's get started!

```


#### Methods 

There are several analyses or tests that must be performed on the data throughout the course of this assignment. We will begin by just reading in and probing the data to see what the status of it is initially. We will then clean the data if necessary (spoiler alert: it was necessary), sort out null values and outliers, view correlations, and create regression models based on variations of parameters in the original data. These variations are determined by looking at correlation values, VIF values, and AIC values, as well as gaining a summary of the model performance after it has been created. 

From all of this work, we will hope to gain a better understanding of how good models are created and what steps can be taken to improve models.

```{r}
# run tests 
# Pick a dataset of your choosing that is appropriate for MLR. (Should be numerical and not have a lot of missing values). Make sure to comment in your report where you downloaded the dataset.
# 
# I'm choosing an air quality dataset from the following link: https://archive.ics.uci.edu/ml/datasets/Air+Quality#
#
# Attributes
# 0 Date (DD/MM/YYYY)
# 1 Time (HH.MM.SS)
# 2 True hourly averaged concentration CO in mg/m^3 (reference analyzer)
# 3 PT08.S1 (tin oxide) hourly averaged sensor response (nominally CO targeted)
# 4 True hourly averaged overall Non Metanic HydroCarbons concentration in microg/m^3 (reference analyzer)
# 5 True hourly averaged Benzene concentration in microg/m^3 (reference analyzer)
# 6 PT08.S2 (titania) hourly averaged sensor response (nominally NMHC targeted)
# 7 True hourly averaged NOx concentration in ppb (reference analyzer)
# 8 PT08.S3 (tungsten oxide) hourly averaged sensor response (nominally NOx targeted)
# 9 True hourly averaged NO2 concentration in microg/m^3 (reference analyzer)
# 10 PT08.S4 (tungsten oxide) hourly averaged sensor response (nominally NO2 targeted)
# 11 PT08.S5 (indium oxide) hourly averaged sensor response (nominally O3 targeted)
# 12 Temperature in °C
# 13 Relative Humidity (%)
# 14 AH Absolute Humidity
#
#
#

# Objective: Find the best multilinear regression model of the dataset of your choosing. ?
# 
# You must include:
#   1. Box plot and histogram of the dependent variable

# First we load the libraries and data
library(data.table)
library(dplyr)
library('ggplot2')
library('corrplot')
library('Rmisc')
library('MASS')
library('car')

# Load data (pick either white or red wine)
dt <- read.csv2("C:\\Users\\jerem\\OneDrive\\Documents\\School\\_REGIS\\2022-05_Summer\\MSDS660\\Week3\\AirQualityUCI\\AirQualityUCI.csv", sep = ";")

# Change data to datatable or dataframe 
dt <- as.data.frame(dt)

### Exploratory ### 
# Look at summary to see if there are any NAs or characters in the dataset. 
# If there are NAs and/or characters remove them.
head(dt)
nrow(dt)
ncol(dt)
str(dt)
summary(dt)

which(is.na(dt$Date))
which(is.na(dt$Time))
which(is.na(dt$CO.GT.))
which(is.na(dt$PT08.S1.CO.))
which(is.na(dt$NMHC.GT.))
which(is.na(dt$C6H6.GT.))
which(is.na(dt$PT08.S2.NMHC.))
which(is.na(dt$N0x.GT.))
which(is.na(dt$PT08.S3.N0x.))
which(is.na(dt$N02.GT.))
which(is.na(dt$PT08.S4.N02.))
which(is.na(dt$PT08.S5.03.))
which(is.na(dt$T))
which(is.na(dt$RH))
which(is.na(dt$AH))
#which(is.na(dt$X))
#which(is.na(dt$X.1))

sum(is.na(dt$PT08.S1.CO.))
sum(is.na(dt$NMHC.GT.))
sum(is.na(dt$PT08.S2.NMHC.))
sum(is.na(dt$X))
sum(is.na(dt$X.1))

# It looks like we need to remove the X and X.1 columns as they contain null values only.

dtnona <- dt[-c(16, 17)]

names(dt)
names(dtnona)
nrow(dtnona)
dtnona <- dtnona[complete.cases(dtnona), ]
names(dtnona)
nrow(dtnona)

# Let's verify we removed all the NAs now

which(is.na(dtnona$Date))
which(is.na(dtnona$Time))
which(is.na(dtnona$CO.GT.))
which(is.na(dtnona$PT08.S1.CO.))
which(is.na(dtnona$NMHC.GT.))
which(is.na(dtnona$C6H6.GT.))
which(is.na(dtnona$PT08.S2.NMHC.))
which(is.na(dtnona$N0x.GT.))
which(is.na(dtnona$PT08.S3.N0x.))
which(is.na(dtnona$N02.GT.))
which(is.na(dtnona$PT08.S4.N02.))
which(is.na(dtnona$PT08.S5.03.))
which(is.na(dtnona$T))
which(is.na(dtnona$RH))
which(is.na(dtnona$AH))

# Okay! so we've cleaned all null values. Now I'm going to remove the date column and convert time to a numerical hour

dtfinal <- dtnona[-c(1)]
names(dtfinal)
head(dtfinal)

dtfinal$Time = as.numeric(substr(dtfinal$Time, start = 1, stop = 2))
head(dtfinal)

summary(dtfinal)
str(dtfinal)

# Awesome, all the data is clean and numeric now

hist(dtfinal$CO.GT., main="Air Quality, CO2 Concentration")
par(new = TRUE)
boxplot(dtfinal$CO.GT., horizontal = TRUE, col = rgb(0, 0.8, 1, alpha = 0.5))
box()

# it looks like we want to remove all the negative concentrations, these are erroneous

dtfinal <- subset(dtfinal, dtfinal$CO.GT. >= 0)

hist(dtfinal$CO.GT., main="Air Quality, CO2 Concentration")
par(new = TRUE)
boxplot(dtfinal$CO.GT., horizontal = TRUE, col = rgb(0, 0.8, 1, alpha = 0.5))
box()

# now let's remove outliers

Q1 <- quantile(dtfinal$CO.GT., .25)
Q3 <- quantile(dtfinal$CO.GT., .75)
IQR <- IQR(dtfinal$CO.GT.)
#only keep rows in dataframe that have values within 1.5*IQR of Q1 and Q3
dtfinal <- subset(dtfinal, dtfinal$CO.GT.> (Q1 - 1.5*IQR) & dtfinal$CO.GT.< (Q3 + 1.5*IQR))

# how are things looking now?

hist(dtfinal$CO.GT., main="Air Quality, Outliers Removed")
par(new = TRUE)
boxplot(dtfinal$CO.GT., horizontal = TRUE, col = rgb(0, 0.8, 1, alpha = 0.5))
box()

# looks good! let's view correlations now

# 2. A correlation plot of all the numerical variables

corrplot(cor(dtfinal), method = 'square')
corrplot(cor(dtfinal), method = 'number')

# it looks like "True hourly averaged concentration CO in mg/m^3 (reference analyzer)" is our dependent variable, CO.GT

#Correlations
#most positive
# N0x.GT. (0.72), PT08.S5.03. (0.65), PT08.S2.NMHC. (0.61), N02.GT. (0.49), PT08.S1.CO. (0.46), PT08.S4.N02. (0.37), Time (0.36)
#most negative
# PT08.S3.N0x. (-0.57)

# 3. A MLR model that has a summary, residual plots, and a VIF analysis

# First, I'm going to create a model that takes every variable

m1 <- lm(CO.GT. ~ Time + PT08.S1.CO. + NMHC.GT. + C6H6.GT. + PT08.S2.NMHC. + NOx.GT. + PT08.S3.NOx. + NO2.GT. + PT08.S4.NO2. + PT08.S5.O3. + T + RH + AH, data = dtfinal)
summary(m1)
par(mfrow = c(2, 2))
plot(m1)
vif_values1 <- vif(m1)
vif_values1
par(mfrow = c(1,1))
barplot(vif_values1, main= "VIF Values m1", horiz = TRUE, col = "steelblue", las=2)
abline(v = 10, lwd = 3, lty = 2)

# Next, I want to create a model that just takes the most highly correlated variables with CO.GT, mentioned above
names(dtfinal)
dtm2 <- dtfinal[ , c('CO.GT.','NOx.GT.', 'PT08.S5.O3.', 'PT08.S2.NMHC.', 'NO2.GT.', 'PT08.S1.CO.', 'PT08.S4.NO2.', 'Time', 'PT08.S3.NOx.')]
names(dtm2)
corrplot(cor(dtm2), method = 'square')
corrplot(cor(dtm2), method = 'number')
m2 <- lm(CO.GT. ~ NO2.GT. + NOx.GT. + PT08.S5.O3. + PT08.S2.NMHC. + PT08.S1.CO. + PT08.S4.NO2. + Time + PT08.S3.NOx., data = dtm2)
summary(m2)
par(mfrow = c(2, 2))
plot(m2)
vif_values2 <- vif(m2)
vif_values2
par(mfrow = c(1,1))
barplot(vif_values2, main= "VIF Values m2", horiz = TRUE, col = "steelblue", las=2)
abline(v = 10, lwd = 3, lty = 2)

# Now based on the VIF value I will remove PT08.S1.CO., PT08.S2.NMHC., and PT08.S5.O3. as they all have VIF values above 10
dtm25 <- dtm2[ , c('CO.GT.','NOx.GT.', 'NO2.GT.', 'PT08.S4.NO2.', 'Time', 'PT08.S3.NOx.')]
names(dtm25)
corrplot(cor(dtm25), method = 'square')
corrplot(cor(dtm25), method = 'number')
m25 <- lm(CO.GT. ~ NO2.GT. + NOx.GT. + PT08.S4.NO2. + Time + PT08.S3.NOx., data = dtm25)
summary(m25)
par(mfrow = c(2, 2))
plot(m25)
vif_values25 <- vif(m25)
vif_values25
par(mfrow = c(1,1))
barplot(vif_values25, main= "VIF Values m25", horiz = TRUE, col = "steelblue", las=2)
abline(v = 10, lwd = 3, lty = 2)


# 4. Comment in the R code a justification on why you removed, combined, or left all variables/observations.

# I chose to only keep values which had a high degree of correlation with CO.GT. (abs>0.3) and which had a VIF value under 10. 
# This led to only choosing 'CO.GT.','NOx.GT.', 'NO2.GT.', 'PT08.S4.NO2.', 'Time', and 'PT08.S3.NOx.'

# 5. Kudos bonus if you create dummy variables from character values.

# This step was not needed as the character values (strings) could be converted to numeric as it was a Time variable

# 6. a stepAIC analysis

fit1 <- stepAIC(m1, direction = "both")
fit1
summary(fit1)

dtm3 = dtfinal[ , c('CO.GT.', 'Time', 'NMHC.GT.', 'PT08.S1.CO.', 'PT08.S2.NMHC.', 'C6H6.GT.', 'NOx.GT.', 'NO2.GT.', 'PT08.S4.NO2.', 'PT08.S5.O3.', 'T', 'RH', 'AH')]
corrplot(cor(dtm3), method = 'square')
corrplot(cor(dtm3), method = 'number')
m3 <- lm(formula = CO.GT. ~ Time + PT08.S1.CO. + NMHC.GT. + C6H6.GT. + PT08.S2.NMHC. + NOx.GT. + NO2.GT. + PT08.S4.NO2. + PT08.S5.O3. + T + RH + AH, data = dtm3)
summary(m3)
par(mfrow = c(2, 2))
plot(m3)
vif_values3 <- vif(m3)
vif_values3
par(mfrow = c(1,1))
barplot(vif_values3, main= "VIF Values m3", horiz = TRUE, col = "steelblue", las=2)
abline(v = 10, lwd = 3, lty = 2)

# based on the correlation values and VIF values, I will drop  from dtm3 to get dtm4 model

dtm4 = dtm3[ , c('CO.GT.', 'Time', 'NOx.GT.', 'NO2.GT.')]
corrplot(cor(dtm4), method = 'square')
corrplot(cor(dtm4), method = 'number')
m4 <- lm(formula = CO.GT. ~ Time + NOx.GT. + NO2.GT., data = dtm4)
summary(m4)
par(mfrow = c(2, 2))
plot(m4)
vif_values4 <- vif(m4)
vif_values4
par(mfrow = c(1,1))
barplot(vif_values4, main= "VIF Values m4", horiz = TRUE, col = "steelblue", las=2)
abline(v = 10, lwd = 3, lty = 2)
```


#### Results 


The results below tell us that m3 had the best model performance of any of the models created. Interestingly enough, the first model created which included all of the parameters, m1, was basically tied for the best performance of the models. I'm not sure why this would be, more insight needs to be gained from the data. These results tell us that the stepAIC analysis was successful in determining what would be the best data parameters to include in the regression model. 

```{r}
# report results 

# 7. Summary, residual plots, and a VIF analysis of the final best model you created.

anova(m1, m2)
# so m2 is better than m1, p < 2.2e-16
anova(m1, m25)
# m25 is better than m1, p < 2.2e-16
anova(m1, m3)
# m1 is better than m3, p=0.7365
anova(m1, m4)
# m4 is better than m1, p < 2.2e-16
anova(m2, m25)
# m25 is better than m2, p < 2.2e-16
anova(m2, m3)
# m3 is better than m2, p < 2.2e-16
anova(m2, m4)
# m4 is better than m2, p < 2.2e-16
anova(m25, m3)
# m3 is better than m25, p < 2.2e-16
anova(m25, m4)
# m4 is better than m25, p < 2.2e-16
anova(m3, m4)
# m4 is better than m3, p < 2.2e-16

# Adjusted R^2
# The bigger adjusted R^2, the better the fit
# m1: 0.8696
# m2: 0.7689
# m25: 0.7493
# m3: 0.8696
# m4: 0.5636

# RSE
# The smaller the RSE, the better the fit
# m1: 0.4343
# m2: 0.578
# m25: 0.6021
# m3: 0.4342
# m4: 0.7944

# So the best model I created was m3, the model suggested by the stepAIC analysis!
summary(m3)
par(mfrow = c(2, 2))
plot(m3)
vif_values3 <- vif(m3)
vif_values3
par(mfrow = c(1,1))
barplot(vif_values3, main= "VIF Values m3", horiz = TRUE, col = "steelblue", las=2)
abline(v = 10, lwd = 3, lty = 2)

# 8. AIC scores of all models built
# the smaller the AIC, the better the fit
AIC(m1)
AIC(m2)
AIC(m25)
AIC(m3)
AIC(m4)






```

#### Conclusion 

This assignment was a bit daunting in the beginning just because a random dataset must be used. This always can create a bit more of an organically difficult experience, such as one which would occur at a job, or some other natural situation. The data cleaning may be a bit more intensive than if the data were already provided for us but I enjoy this additional challenge. The dataset I ended up whoosing was not too dirty and working with it and cleaning it was not overly challenging. I still have a bit of trouble evaluating how to create a good model but I believe I learned a few techniques throughout this assignment to simply help create more models. Then we can evaluate them by looking at AIC score, R-squared (adjusted), and the anova package.

Thank you!
Jeremy Beard