---
title: "week-5-discussion"
output: 
    word_document: default
  pdf_document: default
  html_document:
    df_print: paged
date: "2022-08-03"
---

## MULTI WAY ANOVA IN CLASS ASSIGNMENT ## 

# Load the libraries


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library('data.table')
library('ggpubr')
library('magrittr')
library('dplyr')
```

# Load 'Interactions_Categorical.csv' data set
# Convert data set to data.table
```{r}

dt <- read.csv("/Users/alisha/Downloads/Interactions_Categorical.csv")
dt <- as.data.table(dt)
str(dt)
```
# Check structure of dt
```{r}
dim(dt)
str(dt)
summary(dt)
head(dt)
```
# Convert Food and Condiment to factors
```{r}
cols <- c("Food", "Condiment")

dt %<>% mutate_each_(funs(factor(.)),cols)
str(dt)
```
# Plot histogram of Enjoyment
```{r}
hist(dt$Enjoyment,
     prob=TRUE,
     main = "Histogram of Enjoyment",
     xlab = "Number of interactions",
     ylab = "Density",
     col="Red")
```
# Plot Enjoyment vs the 2 other factors 

```{r plot}
plot.design(Enjoyment ~ ., data = dt)
```
# Plot Individual Boxplots with means

```{r}
boxplot(Enjoyment ~ Food, data = dt, ylab = 'Enjoyment', xlab = 'Food')
points(dt[, mean(Enjoyment), by=Food])
boxplot(Enjoyment ~ Condiment, data = dt, ylab = 'Enjoyment', xlab = 'Condiment')
points(dt[, mean(Enjoyment), by=Condiment])

```
# Create interaction plot looking at Condiment and Food

```{r}
interaction.plot(x.factor = dt$Condiment,
                 trace.factor = dt$Food, 
                 response = dt$Enjoyment,
                 fun = mean, 
                 type = "b",  # shows each point
                 main = "Interaction Plot",
                 legend = TRUE,
                 trace.label = "FOOD",
                 xlab = "Condiment",
                 ylab="Enjoyment",
                 pch=c(1),
                 col = c("Green"))
```
# Build ANOVA model - the * is giving interactions. Show anova fit summary
```{r fit}
fit <- aov(Enjoyment ~ Food * Condiment, data = dt)
summary(fit)
```

# Perform TukeyHSD to check if which interactions have a significant difference
```{r}
TukeyHSD(fit)
```

# Plot the residuals of the fit
```{r}
par(mfrow = c(2,2))
plot(fit)
```

# Perform Shapiro test to see if residuals are normald distributed.
```{r}
shapiro.test(residuals(fit))
hist(residuals(fit))
```
Reference
Frost, J. (2022, July 22). Understanding interaction effects in statistics. Statistics By Jim. Retrieved August 3, 2022, from https://statisticsbyjim.com/regression/interaction-effects/ 

